The function--sheaf dictionary for smooth commutative group schemes over finite fields, with applications over local fields

### 2014-08-20 dictionary.tex
Changes to Applications: QC(Gr(X)
Changes to Applications. Removed dead fragments.
Still working on Applications; getting it ready for our meeting this
afternoon.

\title[The function--sheaf dictionary]{The function--sheaf dictionary for smooth commutative group schemes over finite fields}

@@

Our interest in the function--sheaf dictionary for smooth commutative group schemes over $\Fq$, especially etale commutative group schemes, comes from an application to $p$-adic representation theory.
In \cite{cunningham-roe:13a} we consider quasicharacter sheaves on the
Greenberg transform of the N\'eron model of an algebraic torus over an arbitrary local field.
The ability to work with non-connected group schemes,
of which the component group of the special fibre of the Neron model of a torus is an important example,
plays a crucial role in this application,
since these group schemes are not connected in general.
This process of creating a category from the group of quasicharacters of such a torus
informs our choice of the term quasicharacter sheaf, here.

@@

Note that Proposition~\ref{prop:autornaught_etale} tells us that if $G$ is an etale commutative group scheme over $\Fq$ then every hom-set is either trivial or $\left(G(\bFq)_{\Weil{}}\right)^*$, non-canonically.

@@

In this section we find $\QCiso{G}$ and $\Hom(\qcs{L},\qcs{L}')$ for every smooth commutative group scheme over $\Fq$; note that we found these in Section~\ref{sec:etale} when $G$ is etale.

@@

\begin{proposition}\label{prop:connected}
 If $G$ is a connected, commutative algebraic group over $\Fq$ then:
 \begin{enumerate}
 \item $\TrFrob{G} : \QCiso{G} \to G(\Fq)^*$ is an isomorphism of groups;
 \item every quasicharacter sheaf on $G$ is isomorphic to one defined by a discrete isogeny;%as in Proposition~\ref{prop:finite};
 \item $\Aut(\qcs{L}) = 1$, for all quasicharacter sheaves $\qcs{L}$ on $G$;
  \item every $\rho \in \Hom(\qcs{L},\qcs{L}')$ is either trivial or an isomorphism,
  for all quasicharacter sheaves $\qcs{L}$, $\qcs{L}'$ on $G$.
 \end{enumerate}
 \end{proposition}

@@

We say that a morphism $f : H\to G$ of commutative group $\Fq$-schemes is a {\it discrete isogeny} to $G$ if it is a finite, surjective \'etale morphism and
the action of $\Gal(\bFq/\Fq)$ on the \'etale group scheme $\ker f$ is trivial.

@@

\begin{theorem}\label{thm:autornaught}
Let $G$ be a smooth, commutative group scheme over $\Fq$.
For all quasicharacter sheaves $\qcs{L}$ and $\qcs{L}'$ on $G$,
every $\rho \in \Hom(\qcs{L},\qcs{L}')$ is either trivial or an isomorphism.
Moreover, $\Aut(\qcs{L}) \iso \Hom(\pi_0(\bG)_{\Weil{}},\EEx)$, canonically.
\end{theorem}
\begin{proof} 
For the first part of the proposition, argue as in the analogous part of Proposition~\ref{prop:autornaught_etale}.
For the second, fix $\qcs{L} = (\gqcs{L},\mu,\phi)$ and consider the homomorphism $\Aut(\qcs{L}) \to \Hom(\bG_{\Weil{}},\EEx)$ defined in Proposition~\ref{prop:autornaught_etale}.
This group homomorphism is clearly injective: if $\rho_x = \id_{\qcs{L}_x}$ for every $x\in \bG$ then $\rho = \id_{\qcs{L}}$.
Note that the image of $\Aut(\qcs{L}) \to \Hom(\bG_{\Weil{}},\EEx)$ is contained in $\Hom_\text{cts}(\bG_{\Weil{}},\EEx)$, where $\bG_{\Weil{}}$ is given the topology of the base of the espace étalé attached to $\gqcs{L}$.
Since $\ell$ is invertible in $\Fq$, it follows that $\Hom_\text{cts}(\bG_{\Weil{}},\EEx) = \Hom(\pi_0(\bG_{\Weil{}}),\EEx)$. 
Note that $\pi_0(\bG_{\Weil{}})=\pi_0(\bG)_{\Weil{}}$. 
To see that $\Aut(\qcs{L}) \to \Hom(\pi_0(\bG)_{\Weil{}},\EEx)$ is surjective, begin with $\theta\in\Hom(\pi_0(\bG)_{\Weil{}},\EEx)$ and, for each $[x] \in \pi_0(\bG)_{\Weil{}}$ define $\bar\rho^y: \gqcs{L}^y \to \gqcs{L}^y$ by scalar multiplication by $\theta([x])\in \EEx$ for each $y\in [x]$.
This defines an isomorphism $\bar\rho : \gqcs{L}\to \gqcs{L}$ of local systems on $\bG$ compatible with $\mu$ and $\phi$, and thus an isomorphism $\rho :\qcs{L}\to \qcs{L}$ with maps to $\theta$ under $\Aut(\qcs{L}) \to \Hom(\pi_0(\bG)_{\Weil{}},\EEx)$.
\end{proof}

@@

in this section we use quasicharacter sheaves to geometrize and categorify the group 
of admissible characters of smooth commutative group schemes over local fields that admit Neron models.

Let $R$ complete discrete valuation ring $R$ with residue field $\Fq$.
Write $\m$ for the maximal ideal of $R$ and $K$ for the quotient field for $R$.
Set $R_n = R/\m^{n+1}$ and $S_n = \Spec{R_n}$ and $S = \Spec{R}$; 
note that $S_0 = \Spec{\Fq}$ is the special fibre of $S$.

\subsection{Greenberg transform}

@@

For any $n\in \NN_0$, we will write
\[
\Gr_n : \partop{\text{Schemes}}{\text{lft over $R$}} \to \partop{\text{Schemes}}{\text{lft over $\Fq$}}
\]
for the functor produced by composing pullback along $S_n \to S$ with the Greenberg functor. This functor respects \'etale morphisms, smooth morphisms and geometric components, \todo{open immersions? closed immersion?} and
\[
\Gr_n(X)(\Fq) \iso X(R_n),
\]
canonically, for any $X$ lft over $R$.

For any $n\leq m$,  the surjective ring homomorphism $R_{m} \to R_n$ determines a
natural transformation $\Gr_{m} \to \Gr_n$ between additive functors.
Crucially, $(\Gr_{m}\to \Gr_n)(X)$ is an affine morphism of $\Fq$-schemes, for every $R$-scheme $X$ and every $n\leq m$ \cite{bertrapelle-gonzales:Greenberg}*{Prop 4.3}.
This observation is key to the proof that, for any $X$ lft over $S$, the projective limit 
\[
\Gr(X) \ceq \varprojlim_{n\in \NN_0} \Gr_n(X)
\]
taken with respect to the surjective morphisms $(\Gr_{m}\to \Gr_n)(X)$,
exists in the category of group schemes over $\Fq$.
This leads to the definition of what we call the {\it Greenberg transform}:
\[
\Gr : \partop{\text{Schemes}}{\text{lft over $R$}} \to \partop{\text{Schemes}}{\text{lft over $\Fq$}}.
\]
By construction, the $\Fq$-scheme $\Gr(X)$ comes equipped with morphisms 
\[
P_n(X) : \Gr(X) \to \Gr_n(X),\qquad \forall n\in \NN_0.
\]
The Greenberg transform $\Gr$ respects \'etale morphisms, smooth morphisms  \todo{open immersions? closed immersion?} and geometric components, and
\[
\Gr(X)(\Fq) \iso X(R),
\]
canonically, for any $X$ lft over $R$.

\subsection{Quasicharacter sheaves on the Greenberg transform of a Neron model}

Now suppose $X_K$ is a commutative algebraic group over $K$ 
that admits a Neron model, $X$.
Then $\Gr(X)$ is commutative group scheme over $\Fq$
with component group scheme
$\pi_0(\Gr(X))= \pi_0(X) \times_S S_0$
and
\[
\Gr(X)(\Fq)\iso X_K(K)
\]
canonically.
For every $n\in \NN_0$,
$\Gr_n(X)$ is {\it smooth} commutative group scheme over $\Fq$
with component group scheme
$\pi_0(\Gr_n(X))= \pi_0(X) \times_S S_0$.

A {\it quasicharacter sheaf for $X_K$} is a triple
$\qcs{F}\ceq (\gqcs{F},\mu_\qcs{F},\phi_\qcs{F})$ where
\begin{enumerate}
\item[(0)] $\gqcs{F}$ is a constructible $\ell$-adic sheaf on $\Gr(X)_{\bFq} \ceq \Gr(X)\otimes \bFq$,
\item[(1)]  $\mu_\qcs{F}: \bm^* \gqcs{F} \to \gqcs{F}\boxtimes\gqcs{F}$ is an isomorphism $\ell$-adic sheaves, where $\bm : \Gr(X)_{\bFq} \times \Gr(X)_{\bFq} \to \Gr(X)_{\bFq}$ is multiplication for $\Gr(X)_{\bFq}$,
\item[(2)] $\phi_\qcs{F} : \Frob{\Gr(X)}^* \gqcs{F} \to \gqcs{F}$ is an
  isomorphism of constructible $\ell$-adic sheaves, where $\Frob{\Gr(X)}^* : \Gr(X)_{\bFq}  \to \Gr(X)_{\bFq}$ is Frobenius for $\Gr(X)_{\bFq}$,
\end{enumerate}
such that, for some $n\in \NN_0$,
\begin{enumerate}
\item[(3)]
$\qcs{F} = P_n(X)^*\qcs{L}$ for some quasicharacter sheaf $\qcs{L}$ on $\Gr_n(X)$, where pull-back by $P_n(X)$ is defined as in Proposition~\ref{prop:pullback}.
\end{enumerate}

\begin{theorem}
Let $X_K$ be a commutative algebraic group over $K$ with Neron model $X$ over $R$. 
\begin{enumerate}
\item[(0)] Quasicharacter sheaves for $X_K$ form a rigid monoidal category, $\QC(X_K)$, in a natural way. 
\item[(1)] There is a canonical short exact sequence of abelian groups 
\[
0 \to \Hh^2(\pi_0(X)_{\bFq},\EEx)^{\Frob{}} \to \QCiso{X_K} \to \Hom_\text{ad}(X_K(K),\EEx) \to 0
\] 
where $\Hom_\text{ad}(X_K(K),\EEx)$ is the group of characters of $X_K(K)$ that factor through $X(R)\to X(R_n)$, for some $n\in\NN_0$.
\item[(2)] For all quasicharacter sheaves $\qcs{F}$, $\qcs{F}'$ for $X_K$, and for every $\rho \in \Hom(\qcs{F},\qcs{F}')$, either $\rho$ is trivial or $\rho$ is an isomorphism.
\item[(3)] For all quasicharacter sheaves $\qcs{F}$ for $X_K$,
\[
\Aut(\qcs{L}) \iso \Hom((\pi_0(X)_{\Fq})_{\Weil{}},\EEx),
\]
canonically.
\end{enumerate}
\end{theorem}

@@

\section{Fragments}

\subsection{Finite and bounded - fragments}

On general smooth group $\Fq$-schemes $G$, 
there are bounded quasicharacter sheaves which are not finite.  
Identifying $\fQC(G)$ with its image under $B_G \circ L_G$, as defined in
Proposition~\ref{prop:BG} and in the proof of Proposition~\ref{prop:bounded},
we will see in Corollary~\ref{cor:bounded-and-finite} that
a quasicharacter sheaf is finite if and only if its Trace of Frobenius has finite image.

\begin{corollary} \label{cor:bounded-and-finite}
Let $G$ be a smooth commutative group scheme over $\Fq$.
\begin{enumerate}
\item $\qcs{L} \in \QC(G)$ is bounded if and only if $\trFrob{\qcs{L}}$ has bounded image,
\item $\qcs{L}\in \QC(G)$ is finite if and only if $\trFrob{\qcs{L}}$ has finite image.
\end{enumerate}
\end{corollary}

\begin{proof}
We first reduce to the case that $G$ is etale.
On the connected component
$G^0$, all quasicharacter sheaves are bounded and finite by Proposition~\ref{prop:connected}.
On the other hand, $G^0(\Fq)$ is finite by Lemma~\ref{lem:G0alg-grp} and thus Trace of Frobenius
has finite and therefore bounded image.  If $\chi \in G(\Fq)^*$ then there is some finite-image character $\chi_0$
with the same restriction to $G^0(\Fq)$ since $G^0(\Fq)$ is lies inside the torsion part of
the finitely generated abelian group $G(\Fq)$.  Therefore $\chi$ is bounded (resp. has finite image)
if and only if $\chi \cdot \chi_0^{-1}$ is bounded (resp. has finite image).  But $\chi \cdot \chi_0^{-1}$ descends
to a character of $\pi_0(G)$.  A similar argument relates boundedness/finiteness of quasicharacter
sheaves on $G$ to those on $\pi_0(G)$, using the fact that the preimage constructed in the proof of
Proposition~\ref{prop:restriction} is a finite quasicharacter sheaf on $G$.
Using these equivalences, it is enough to prove Corollary~\ref{cor:bounded-and-finite} when $G$ is etale.

In the case that $G$ is \'etale, which we now suppose,
one can easily detect when a quasicharacter sheaf is bounded or finite.
With reference to Section~\ref{sec:S},
let $S_G(\qcs{L}) = [\alpha \oplus \beta]$
and let $b : {\bar G}\to \EEx$ be the multiplicative version of $\beta(\Frob{})$.
Then $\qcs{L}$ is finite if and only if
$b(x)$ is a root of unity for all $x\in {\bar G}$.
By \cite{beilinson-bernstein-deligne:81a}*{Rem. 5.2.1},
$\qcs{L}$ is bounded if and only if $b(x) \in \Zlx$ for all $x\in {\bar G}$.
\end{proof}

>>>>

\title[The function--sheaf dictionary, with applications over local fields]{The function--sheaf dictionary for smooth commutative group schemes over finite fields, with applications over local fields}

@@

Our interest in the function--sheaf dictionary for smooth commutative group schemes over $\Fq$, especially etale commutative group schemes, comes from an application to $p$-adic representation theory.
In \cite{cunningham-roe:13a} we consider quasicharacter sheaves on the
Greenberg transform of the N\'eron model of an algebraic torus over an arbitrary local field.
The ability to work with non-connected group schemes,
of which the component group of the special fibre of the Néron model of a torus is an important example,
plays a crucial role in this application,
since these group schemes are not connected in general.
This process of creating a category from the group of quasicharacters of such a torus
informs our choice of the term quasicharacter sheaf, here.

@@

%Note that Proposition~\ref{prop:autornaught_etale} tells us that if $G$ is an etale commutative group scheme over $\Fq$ then every hom-set is either trivial or $\{ 0 \} \cup \left(G(\bFq)_{\Weil{}}\right)^*$, non-canonically.

@@

%In this section we find $\QCiso{G}$ and $\Aut(\qcs{L})$ for commutative algebraic group over $\Fq$; note that we found these in Section~\ref{sec:etale} when $G$ is etale.

@@

\begin{proposition}\label{prop:connected}
 If $G$ is a connected, commutative algebraic group over $\Fq$ then:
 \begin{enumerate}
 \item $\TrFrob{G} : \QCiso{G} \to G(\Fq)^*$ is an isomorphism of groups;
 \item every quasicharacter sheaf on $G$ is isomorphic to one defined by a discrete isogeny;%as in Proposition~\ref{prop:finite};
 \item $\Aut(\qcs{L}) = 1$, for all quasicharacter sheaves $\qcs{L}$ on $G$;
  \item for all quasicharacter sheaves $\qcs{L}$, $\qcs{L}'$ on $G$ and for every $\rho \in \Hom(\qcs{L},\qcs{L}')$, either $\rho$ is trivial or an isomorphism.
 \end{enumerate}
 \end{proposition}

@@

%We say that a morphism $f : H\to G$ of commutative group $\Fq$-schemes is a {\it discrete isogeny} to $G$ if it is a finite, surjective \'etale morphism and
the action of $\Gal(\bFq/\Fq)$ on the \'etale group scheme $\ker f$ is trivial.

@@

\begin{theorem}\label{thm:autornaught}
Let $G$ be a smooth commutative group scheme over $\Fq$.
For all quasicharacter sheaves $\qcs{L}$ and $\qcs{L}'$ on $G$,
every $\rho \in \Hom(\qcs{L},\qcs{L}')$ is either trivial or an isomorphism.
Moreover, $\Aut(\qcs{L}) \iso \Hom(\pi_0(\bG)_{\Weil{}},\EEx)$, canonically.
\end{theorem}
\begin{proof} 
For the first part of the proposition, argue as in the analogous part of Proposition~\ref{prop:autornaught_etale}.
For the second, fix $\qcs{L} = (\gqcs{L},\mu,\phi)$ and consider the homomorphism $\Aut(\qcs{L}) \to \Hom(\bG_{\Weil{}},\EEx)$ defined in Proposition~\ref{prop:autornaught_etale}.
This group homomorphism is clearly injective: if $\rho_x = \id_{\qcs{L}_x}$ for every $x\in \bG$ then $\rho = \id_{\qcs{L}}$.
Note that the image of $\Aut(\qcs{L}) \to \Hom(\bG_{\Weil{}},\EEx)$ is contained in $\Hom_\text{cts}(\bG_{\Weil{}},\EEx)$, where $\bG_{\Weil{}}$ is given the topology of the base of the espace étalé attached to $\gqcs{L}$.
Since $\ell$ is invertible in $\Fq$, it follows that $\Hom_\text{cts}(\bG_{\Weil{}},\EEx) = \Hom(\pi_0(\bG_{\Weil{}}),\EEx)$. 
Note that $\pi_0(\bG_{\Weil{}})=\pi_0(\bG)_{\Weil{}}$. 
To see that $\Aut(\qcs{L}) \to \Hom(\pi_0(\bG)_{\Weil{}},\EEx)$ is surjective, begin with $\theta\in\Hom(\pi_0(\bG)_{\Weil{}},\EEx)$ and, for each $[x] \in \pi_0(\bG)_{\Weil{}}$ define $\bar\rho^y: \gqcs{L}^y \to \gqcs{L}^y$ by scalar multiplication by $\theta([x])\in \EEx$ for each $y\in [x]$.
This defines an isomorphism $\bar\rho : \gqcs{L}\to \gqcs{L}$ of local systems on $\bG$ compatible with $\mu$ and $\phi$, and thus an isomorphism $\rho :\qcs{L}\to \qcs{L}$ with maps to $\theta$ under $\Aut(\qcs{L}) \to \Hom(\pi_0(\bG)_{\Weil{}},\EEx)$.
\end{proof}

@@

In this section we use quasicharacter sheaves to geometrize and categorify the group 
of admissible characters of smooth commutative group schemes over local fields that admit Néron models.

Let $R$ be a complete discrete valuation ring $R$ with residue field $\Fq$.
Write $\m$ for the maximal ideal of $R$ and $K$ for the quotient field for $R$.
Set $S = \Spec{R}$.
For every non-negative integer $n$, set $R_n = R/\m^{n+1}$ and $S_n = \Spec{R_n}$;  
note that $S_0 = \Spec{\Fq}$ is the special fibre of $S$.

\subsection{Review of the Greenberg transform}

@@

For any $n\in \NN_0$, we will write
\[
\Gr^R_n : \partop{\text{Schemes}}{\text{lft over $R$}} \to \partop{\text{Schemes}}{\text{lft over $\Fq$}}
\]
for the functor produced by composing pullback along $S_n \to S$ with the Greenberg functor. 
This functor respects open immersions, closed immersions, \'etale morphisms, smooth morphisms and geometric components, and
\[
\Gr^R_n(X)(\Fq) \iso X(R_n),
\]
canonically, for any $X$ lft over $R$.

For any $n\leq m$,  the surjective ring homomorphism $R_{m} \to R_n$ determines a
natural transformation $P^R_{n\leq m} : \Gr^R_{m} \to \Gr^R_n$ between additive functors.
Crucially, $P^R_{n\leq m}(X): \Gr^R_{m}(X)\to \Gr^R_n(X)$ is an affine morphism of $\Fq$-schemes, for every $R$-scheme $X$ and every $n\leq m$ \cite{bertrapelle-gonzales:Greenberg}*{Prop 4.3}.
This observation is key to the proof that, for any $X$ lft over $S$, the projective limit 
\[
\Gr_R(X) \ceq \varprojlim_{n\in \NN_0} \Gr^R_n(X)
\]
taken with respect to the surjective morphisms $P^R_{n\leq m}(X) : \Gr^R_{m}(X) \to \Gr^R_n(X)$,
exists in the category of group schemes over $\Fq$.
This leads to the definition of what we shall call the {\it Greenberg transform}:
\[
\Gr_R : \partop{\text{Schemes}}{\text{lft over $R$}} \to \partop{\text{Schemes}}{\text{lft over $\Fq$}}.
\]
By construction, the $\Fq$-scheme $\Gr_R(X)$ comes equipped with morphisms 
\[
P^R_n(X) : \Gr^R(X) \to \Gr^R_n(X),\qquad \forall n\in \NN_0.
\]
The Greenberg transform respects open immersions, closed immersions, \'etale morphisms, smooth morphisms and geometric components, and
\[
\Gr_R(X)(\Fq) \iso X(R),
\]
canonically, for any $X$ lft over $R$.

\subsection{Quasicharacter sheaves on the Greenberg transform of a Néron model}

Let $G$ be a commutative group scheme over $\Fq$.
Consider the category $\mathcal{C}(G)$ of triples
$\qcs{F}\ceq (\gqcs{F},\mu_\qcs{F},\phi_\qcs{F})$ where
\begin{enumerate}
\labitem{(C.0)}{C.0} $\gqcs{F}$ is a rank one $\ell$-adic local system on $\bG\ceq G \times_\Spec{\Fq} \Spec{\bFq}$,
\labitem{(C.1)}{C.1} $\mu_\qcs{F}: \bm^* \gqcs{F} \to \gqcs{F}\boxtimes\gqcs{F}$ is an isomorphism $\ell$-adic sheaves, where $\bm : \bG \times \bG \to \bG$ is multiplication $m$ for $G$ extended to $\bG$.
\labitem{(C.2)}{C.2} $\phi_\qcs{F} : \Frob{G}^* \gqcs{F} \to \gqcs{F}$ is an isomorphism of constructible $\ell$-adic sheaves, where $\Frob{G}^* : \bG  \to \bG$ is Frobenius,
\end{enumerate}
subject to the conditions as in \ref{QC.1} and \ref{QC.2},
with morphisms defined as in \ref{QC.3}.
This category $\mathcal{C}(G)$ is rigid monoidal, with tensor product defined as in Definition~\ref{def:QC}. Note that $\mathcal{C}(G) = \QC(G)$ if $G$ is also smooth.

Now suppose $X_K$ is a connected commutative algebraic group over $K$ that admits a Néron model $X$ over $R$.

A {\it quasicharacter of $X_K(K)$} is a group homomorphism $X_K(K) \to \EEx$ that factors through $X_K(X) = X(R)\to X(R_n)$, for some $n\in\NN_0$;
the group of quasicharacters of $X_K(K)$ will be denoted by $\Hom_\text{qc}(X_K(K),\EEx)$.

Then $\Gr_R(X)$ is commutative group scheme over $\Fq$
with component group scheme
\[
\pi_0(\Gr_R(X))= \pi_0(X) \times_S S_0
\]
and $\Fq$-rational points
\[
\Gr_R(X)(\Fq)\iso X_K(K),
\]
canonically.
For every $n\in \NN_0$,
$\Gr^R_n(X)$ is {\it smooth} commutative group scheme over $\Fq$
with component group scheme
$\pi_0(\Gr^R_n(X))= \pi_0(X) \times_S S_0$.

For every non-negative integer $n$, the morphism of $\Fq$-schemes $P^R_n(X) : \Gr_R(X) \to \Gr^R_n(N)$ induces a functor
\[
P^R_n(X)^* : \QC(\Gr^R_n(X)) \to \mathcal{C}(\Gr_R(X)),
\]
as in Proposition~\ref{prop:pullback}.
This functor is full and faithful, so we may view $\QC(\Gr^R_n(X))$ as a full subcategory of $\mathcal{C}(\Gr_R(X))$.

\begin{definition}
A {\it quasicharacter sheaf on $\Gr_R(X)$} is an object $\qcs{F}$ in $\mathcal{C}(\Gr_R(X))$ such that $\qcs{F} \iso P^R_n(X)^*\qcs{L}$ for some quasicharacter sheaf $\qcs{L}$ on $\Gr^R_n(X)$.
Quasicharacter sheaves on $\Gr_R(X)$ form a rigid monoidal category in the obvious way, denoted by $\QC(\Gr_R(X))$.
\end{definition}

Note that we may informally regard $\QC(\Gr_R(X))$ a `union' of the full subcategories $\QC(\Gr^R_n(X))$ in $\mathcal{C}(\Gr_R(X))$.

\begin{theorem}\label{thm:QCX}
Let $X_K$ be a commutative algebraic group over $K$ with Néron model $X$ over $R$. 
\begin{enumerate}
\item[(1)] There is a canonical short exact sequence of abelian groups 
\[
0 \to \Hh^2(\pi_0(X)_{\bFq},\EEx)^{\Frob{}} \to \QCiso{\Gr_R(X)} \to \Hom_\text{qc}(X_K(K),\EEx) \to 0.
\] 
\item[(2)] For all quasicharacter sheaves $\qcs{F}$, $\qcs{F}'$ on $\Gr_R(X)$, and for every $\rho \in \Hom(\qcs{F},\qcs{F}')$, either $\rho$ is trivial or $\rho$ is an isomorphism.
\item[(3)] For all quasicharacter sheaves $\qcs{F}$ on $\Gr_R(X)$,
\[
\Aut(\qcs{F}) \iso \Hom((\pi_0(X)_{\Fq})_{\Weil{}},\EEx),
\]
canonically.
\end{enumerate}
\end{theorem}

\begin{proof}
To begin, fix $n\in \NN$.
The following statements follow directly from Theorems~\ref{thm:snake} and \ref{thm:autornaught} together with our observations about the component group of $\Gr^R_n(X)$.
\begin{enumerate}
\item[(1)] There is a canonical short exact sequence of abelian groups 
\[
0 \to \Hh^2(\pi_0(X)_{\bFq},\EEx)^{\Frob{}} \to \QCiso{\Gr^R_n(X)} \to \Hom(X(R_n),\EEx) \to 0.
\] 
\item[(2)] For all quasicharacter sheaves $\qcs{L}$, $\qcs{L}'$ on $\Gr^R_n(X)$ and for every $\rho \in \Hom(\qcs{L},\qcs{L}')$, either $\rho$ is trivial or $\rho$ is an isomorphism.
\item[(3)] For all quasicharacter sheaves $\qcs{L}$ on $\Gr^R_n(X)$,
\[
\Aut(\qcs{L}) \iso \Hom((\pi_0(X)_{\Fq})_{\Weil{}},\EEx),
\]
canonically.
\end{enumerate}
So Theorem~\ref{thm:QCX} now follows from the fact that each $\QC(\Gr^R_n(X))$ is a full subcategory of $\QC(\Gr_R(X))$ and the observation that $\QC(\Gr_R(X))$ a `union' of the subcategories $\QC(\Gr^R_n(X))$ in $\mathcal{C}(\Gr_R(X))$.
\end{proof}

\subsection{Weil restriction and quasicharacters}\label{sec:wrK}

Let $K'/K$ be a finite Galois extension of local fields and
let $k'/k$ be the corresponding extension of residue fields.
Let $R$ and $R'$ be the rings of integers of $K$ and $K'$, respectively.
Let $X_K$ be a connected commutative algebraic group over $K$
that admits a Néron model $X$ over $R$.
Set $X_{K'} \ceq X_K \times_\Spec{K} \Spec{K'}$
and $X' \ceq X \times_\Spec{R} \Spec{R'}$.

\begin{proposition}\label{prop:wrK}
The canonical closed immersion 
\[
X_K \hookrightarrow \Res_{K'/K} X_{K'}
\]
of $K$-group schemes
induces a map of $\Fq$-group schemes 
\[
f : \Gr_R(X) \to \Res_{k'/k} \Gr_{R'}(X')
\] 
such that the functor
\[
f^* : \QC(\Gr_R(X)) \to \QC(\Res_{k'/k} \Gr_{R'}(X'))
\]
induces
\[
\begin{tikzcd}[column sep=60]
\Hom_{\text{qc}}(X_K(K'), \EEx) \arrow{r}{\chi \mapsto \chi\vert_{X_K(K)}} &\Hom_{\text{qc}}(X_K(K), \EEx).
\end{tikzcd}
\]
\end{proposition}

\begin{proof}
Adapting the argument in \cite{bosch-lutkebohmert-reynaud:NeronModels}*{\S 7.6, Prop 6} to include locally finite type Néron models,
one may show that $\Res_{R'/R}(X')$ is a Néron model for $\Res_{K'/K}(X_{K'})$.
By the Néron mapping property, the canonical closed immersion
\[
X_K\hookrightarrow \Res_{K'/K}(X_{K'})
\]
 extends uniquely to a morphism
\[
X\to \Res_{R'/R}(X')
\]
 of smooth $R$-group schemes.

Applying the functor $\Gr_{n}$ to this morphism
and using \cite{bertrapelle-gonzales:Greenberg}*{Thm 1.1} defines the morphism of smooth group schemes
\begin{equation}\label{men}
f_n: \Gr_{n-1}^R(X) \to \Res_{k'/k} \Gr_{en-1}^{R'}(X')
\end{equation}
Since 
\[
\left(\Res_{k'/k} \Gr_{en-1}^{R'}(X') \right)(\Fq) = \left(\Gr_{en-1}^{R'}(X')\right)(k'),
\]
the pullback functor 
\[
f_n^* : \QC(\Res_{k'/k} \Gr_{en-1}^{R'}(X'))\to \QC(\Gr_{n-1}^{R}(X))
\]
obtained by applying Proposition~\ref{prop:pullback} to \eqref{men} gives
\[
\Hom(X(R'_{en}),\EEx) \to \Hom(X(R_n),\EEx).
\]
Passing to the projective limit, the family of morphisms \eqref{men} determines a morphism of group schemes
\[
f: \Gr_R(X) \to \Res_{k'/k} \Gr_{R'}(X').
\]
which induces the functor
\[
f^* : \QC(\Res_{k'/k} \Gr_{R'}(X'))\to \QC(\Gr_{R}(X))
\]
inducing
\[
\Hom_{\text{qc}}(T(K'),\EEx) \to \Hom_{\text{qc}}(T(K),\EEx).
\]
Argue as in Proposition~\ref{prop:csbe} to see that this is indeed restriction of characters.
\end{proof}

\subsection{Transfer of quasicharacters}\label{sec:transfer}

Let $K$ and $K'$ be local fields with rings of integers $R$ and $R'$ and residue fields $k$ and $k'$, respectively. Set $S = \Spec(S)$ and $S_n = \Spec(R_n)$; likewise define $S' = \Spec{R'}$ and $S'_{n'} = \Spec{R'_{n'}}$.

Let $X_K$ and $X'_{K'}$ be connected commutative algebraic groups over $K$ and $K'$, respectively, that admit Néron models $X$ and $X'$ over $R$ and $R'$, respectively.
Suppose $m$ and $m'$ are non-negative integers such that 
\begin{equation}\label{eq:schematic_transfer}
R_m \iso R'_{m'}
\qquad
\text{and}
\qquad
X \times_{S} S_m \iso X' \times_{S'} S'_{m'}.
\end{equation}
as smooth group schemes over $R_m$. 
Then
$
\Gr^R_m(X) \iso \Gr^{R'}_{m'}(X')
$
as smooth group schemes over $\Fq$ and, by Proposition~\ref{prop:pullback} the isomorphism above determines an equivalence of categories
\begin{equation}\label{eq:categorical_transfer}
\QC(\Gr^R_m(X)) \cong \QC(\Gr^{R'}_{m'}(X'))
\end{equation}
which induces a group isomorphism
$
\QCiso{\Gr^R_m(X)} \cong \QCiso{\Gr^{R'}_{m'}(X')}
$
and thence, by trace of Frobenius, an isomorphism
\begin{equation}\label{eq:transfer}
\Hom(X(R_m),\EEx) \iso \Hom(X'(R'_{m'}),\EEx).
\end{equation}
Recall that $\Hom(X(R_m),\EEx)$ is a subgroup of $\Hom_\text{qc}(X_K(K),\EEx)$ and $\Hom(X'(R'_{m'}),\EEx)$ is a subgroup of $\Hom_\text{qc}(X_K(K),\EEx)$.
The isomorphism \eqref{eq:transfer} is an instance of {\it transfer} of certain quasicharacters between $X_K(K)$ and $X_{K'}(K')$.

Such a relationship can exist between quasicharacters of algebraic tori over local fields, even when the characteristics of $K$ and $K'$ differ
\citelist{\cite{chai-yu:01a} \cite{deligne:84a}*{\S 1}}.

Suppose $T$ and $T'$ are tori over non-archimedian local fields $K$ and $K'$,
splitting over $L$ and $L'$ respectively.
Then $T$ and $T'$ are said to be \emph{$N$-congruent} \cite{chai-yu:01a}*{\S 2} if there are isomorphisms
 \begin{align*}
  \alpha : \OL/\pi_K^N \OL &\to \OO{L'}/\pi_{K'}^N \OO{L'} \\
  \beta : \Gal(L/K) &\to \Gal(L'/K') \\
  \phi : X^*(T) &\to X^*(T')
 \end{align*}
 satisfying the conditions
 \begin{enumerate}
  \item $\alpha$ induces an isomorphism $\OK/\pi_K^N \OK \to \OO{K'}/\pi_{K'}^N \OO{K'}$,
  \item $\alpha$ is $\Gal(L/K)$-equivariant relative to $\beta$, and
  \item $\phi$ is $\Gal(L/K)$-equivariant relative to $\beta$.
 \end{enumerate}
If $T$ and $T'$ are $N$-congruent then $\alpha$, $\beta$ and $\phi$ determine an isomorphism
\begin{equation}\label{transfer}
  \Hom_N(T(K), \EEx) \iso \Hom_N(T'(K'),\EEx).
\end{equation}
Note that if $T$ and $T'$ are $N$-congruent, then they are $N'$ congruent for every
$N' \leq N$, the isomorphism \eqref{transfer} is depth-preserving at integers less than $N$.

One of the main results of \cite{chai-yu:01a} gives an isomorphism of group schemes between the
$\TT_n$ and $\TT'_n$ assuming that $T$ and $T'$ are sufficiently congruent.
They define a quantity $h$ as the smallest integer so that $\pi^h$ lies in the
Jacobian ideal associated to a natural embedding of $T$ into an induced torus \cite{chai-yu:01a}*{\S 8.1}.
Then it follows from \cite{chai-yu:01a}*{Thm. 8.5} that if $N > 3h$ and $T$ and $T'$ are $N$-congruent
then there is a canonical isomorphism of smooth group schemes
 $
  \TT_{N-3h} \to \TT'_{N-3h}
 $
 determined by $\alpha, \beta$ and $\phi$.
Combining this with Theorem~\ref{thm:application} gives the geometrization and categorification of the transfer of characters.

\begin{proposition}\label{prop:transfer}
 With notation above, suppose that $T$ and $T'$ are $N$-congruent and $N > 3h$.  Set $n \ceq N-3h$.
 Then there is a canonical equivalence of categories
 \[
  \QC(\GN{T}'_{n}) \to \QC(\GN{T}_{n})
 \]
 determined by $\alpha, \beta$ and $\phi$ inducing the isomorphism
 \[
  \Hom_{<n}(T'(K'), \EEx) \to \Hom_{<n}(T(K), \EEx)
 \]
 through the trace of Frobenius.
\end{proposition}

\subsection{Algebraic tori and Abelian varieties over local fields}

\begin{corollary}\label{cor:application}
For every algebraic torus $T$ over $K$, the trace of Frobenius
  \[
\trFrob{T}:  \QCiso{\GN{T}} \to \Hom_{\text{qc}}(T(K),\EEx)
  \]
is an isomomorphism.
Moreover, the isomorphism $\trFrob{T}$ is a natural transformation
between the additive functors
\begin{enumerate}
\item[$F_{\QC}$:] $T \mapsto \QCiso{\GN{T}}$
\item[$F_{\text{qc}}$:] $T \mapsto \Hom_{\text{qc}}(T(K),\EEx)$
\end{enumerate}
from the additive category of algebraic tori over $K$ to the category of abelian groups.
\end{corollary}

@@
